--- 
title: 'Online appendix for the paper: "Bayesian Paired-Comparison with the bpcs package"'
author: "David Issa Mattos and Érika Martins Silva Ramos"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: 'This is the online appendix for the paper: "Bayesian Paired-Comparison with the bpcs package."'
always_allow_html: true
---

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
knitr::opts_chunk$set(cache=TRUE)

library(bpcs)
library(tidyverse)
library(knitr)
options(mc.cores = parallel::detectCores())
rstan::rstan_options(auto_write = TRUE)
set.seed(99)
```

# Foreword

This is the online appendix for the paper *"Bayesian Paired-Comparison with the bpcs package"*. It contains a commented and reproducible code for all the analysis, tables and plots presented in the paper.

In the beginning of each study, we show a few lines of the original datasets. These datasets are either available through a link in the paper or through asking the authors directly. Therefore we do not provide or distribute the data in this appendix (or in the code repository).

## Installation of the bpcs package

This appendix was compiled with the version 1.1.0 of the bpcs package.

The development and latest version of the bpcs package can be installed directly from Github with:

```{r eval=F}
remotes::install_github('davidissamattos/bpcs')
```

The latest CRAN release can be installed with:

```{r eval=FALSE}
install.packages('bpcs')
```


We emphasize here that the CRAN version might some versions behind the latest stable release in Github.

## Session info

This appendix is compiled automatically and the following session information was used to generate this appendix:

```{r}
sessionInfo()
```




